\documentclass[12pt,answers]{exam}

\usepackage{amsmath,amsfonts,amssymb,mathtools,physics,commath}
\usepackage{enumitem}
\usepackage{cancel}
\usepackage{todonotes}
\usepackage{outlines}
\newcommand{\vect}[1]{\left\langle #1\right\rangle}

\pagestyle{headandfoot}
\firstpageheadrule
\runningheadrule
\firstpageheader{Math 222}{Exam 2 Version B|Solutions, Page \thepage\ of \numpages}{October 18, 2018}
\runningheader{Math 222}{Exam 2 Version B|Solutions, Page \thepage\ of \numpages}{October 18, 2018}
\runningfooter{}{}{}

\title{2018 Fall Calc 3 Exam 2 Version B|Solutions}
\author{Winston Cheong}
\date{}

\begin{document}
% \maketitle
\begin{questions}
	\question[16]
	Computation
	\begin{parts}
		\part
		Let $f(x, y) = x^3 \exp(2x^2+3xy + y^2)$. Find all of the first partial derivatives. (In case you haven't seen it before, ``$\exp(u)$'' is the same thing as $e^u$.)
		\begin{solution}
			\begin{align*}
				f_x &= 3x^2 \exp(2x^2+3xy + y^2)+ x^3 \exp(2x^2+3xy + y^2) (4x + 3y) \\ 
				f_y &= x^3 \exp(2x^2+3xy + y^2) (3x + 2y)
			\end{align*}
		\end{solution}

		\part
		Let $g(x,y) = \dfrac{x}{\sqrt{4x^2 + y^2}}$. Find the first partial derivative with respect to $x$ and simplify it.
		\begin{solution}
			\begin{align*}
				g_x &= \frac{\sqrt{4x^2+y^2} \cdot 1 - x \cdot \frac{1}{2\sqrt{4x^2+y^2}}\cdot 8x}{4x^2+y^2} \\ 
				&= \frac{4x^2 + y^2 - 4x^2}{(4x^2+y^2)^{3/2}} \\ 
				&= \boxed{\frac{y^2}{(4x^2+y^2)^{3/2}}}
			\end{align*}
		\end{solution}
	\end{parts}

	\newpage
	\question[12]
	A certain differentiable function satisfies:
	\begin{enumerate}[label=(\alph*)]
		\item $f(2,5) = -3$, and $f(-4, 1) = \pi$.
		\item $\nabla f(2,5) = (-4,7)$, and $\nabla f(-4, 1) = (\sqrt 6, e^{-2})$.
	\end{enumerate}
	At each of the two points in question (i.e.~at $(2,5)$ and at $(-4, 1)$) answer the following questions:
	\begin{parts}
		\part In what direction is the function increasing the fastest and what is the rate of change in that direction?
		\begin{solution}~\\
			At $(2, 5)$, the function is increasing the fastest in the direction $\nabla f(2, 5) = \vect{-4,7}$, with rate of change $\|\vect{-4,7}\| = \sqrt{16+49} = \sqrt{65}$.

			At $(-4,1)$, the function is increasing the fastest in the direction $\nabla f(-4,1) = \vect{\sqrt 6, e^{-2}}$, with rate of change $\|\vect{\sqrt 6, e^{-2}}\| = \sqrt{6 + e^{-4}}$.
		\end{solution}

		\part What is the directional derivative in the direction of the vector $\vect{4,-3}$?
		\begin{solution}
			The unit vector in the direction $\vect{4,-3}$ is $\vect{4/5,-3/5}$.
			So
			\begin{align*}
				D_{\vect{4/5,-3/5}} f(2,5) 
				&= \nabla f(2,5) \vdot \vect{4,-3} \cdot \frac15 \\ 
				&= \vect{-4,7} \vdot \vect{4,-3} \cdot \frac15 \\
				&= (-16-21) \cdot \frac15
				= \boxed{\frac{-37}{5}}
				\shortintertext{and}
				D_{\vect{4/5,-3/5}} f(-4,1) 
				&= \nabla f(-4,1) \vdot \vect{4,-3} \cdot \frac15 \\ 
				&= \vect{\sqrt6, e^{-2}} \vdot \vect{4,-3} \cdot \frac15 \\
				&= \boxed{(4 \sqrt 6 - 3 e^{-2}) \cdot \frac15}
			\end{align*}
		\end{solution}

		\part What is the tangent plane and/or the linear approximation at each of the two points?
		\begin{solution}
			\begin{align*}
				(2,5) &:\quad z = -3 + -4(x-2) + 7(y-5) \\ 
				(-4,1) &:\quad z = \pi + \sqrt 6 (x+4) + e^{-2} (y-1)
			\end{align*}
		\end{solution}
	\end{parts}

	\newpage
	\question[12]
	Set up \textbf{but do not solve} the following problems. As part of setting these problems up, you should list the unknowns and the equations that you would need to use to find them. You \textbf{should also do} all of the \textbf{derivative} calculations, but the \textbf{algebra} is totally unmanageable, so do \textbf{not} attempt it!
	\begin{parts}
		\part Maximize $f(x, y) = x^2 \cos(y)$ \\ 
		Subject to $g(x,y) = x^6 + y^6 = 64$.
		\begin{solution}
			\begin{align*}
				\nabla f &= \lambda \nabla g \\ 
				\vect{2x\cos y, -x^2 \sin y}  &= \lambda \vect{6x^5, 6y^5}
			\end{align*}
			The system to solve is:
			\[
				\left\{
					\begin{aligned}
						2x\cos y &= 6\lambda x^5 \\ 
						-x^2 \sin y &= 6\lambda y^5 \\
						x^6 + y^6 &= 64
					\end{aligned}
				\right.
			\]
		\end{solution}

		\part Maximize $F(x,y,z) = \cos(xy^2z^3)$\\
		Subject to $G(x,y,z) = x + 2y + 3z = 0$\\
		and $H(x,y,z) = x^2 + z^2 = 25$.
		\begin{solution}
			\begin{align*}
				\nabla F &= \vect{-y^2z^3\sin(xy^2z^3), -2xyz^3\sin(xy^2z^3), -3xy^2z^2\sin(xy^2z^3)} \\
				\nabla G &= \vect{1,2,3} \\
				\nabla H &= \vect{2x,0,2z}
			\end{align*}
			Setting $\nabla F = \lambda\nabla G + \mu\nabla H$ gives the system to solve:
			\[
				\left\{
					\begin{aligned}
						-y^2z^3\sin(xy^2z^3) &= \lambda + 2\mu x \\
						-2xyz^3\sin(xy^2z^3) &= 2\lambda \\
						-3xy^2z^2\sin(xy^2z^3) &= 3\lambda + 2\mu z \\
						x+2y+3z &= 0 \\
						x^2 + z^2 &= 25
					\end{aligned}
				\right.
			\]
		\end{solution}
	\end{parts}

	\newpage
	\question[14]
	For the function $f(x,y) = x^3 + xy - y^2$ find and classify all of the critical points.
	\begin{solution}
		\begin{align*}
			\nabla f = 0
			&\implies
			\left\{
				\begin{aligned}
					3x^2 + y &= 0\\ 
					x - 2y &= 0
				\end{aligned}
			\right.
		\end{align*}
		Solving this system gives two critical points: 
		$(0,0)$ and $(-\frac1{6}, -\frac1{12})$.
		To classify them, the discriminant is
		\[
			f_{xx} f_{yy} - (f_{xy})^2 
			= 6x(-2) - (1)^2 = -12x - 1
		\]
		For $(0,0)$, the discriminant gives $-1$, so $(0,0)$ is a saddle point.

		For $(-\frac16, -\frac1{12})$, the discriminant gives 1, and $f_{yy} = -2 < 0$, so $(-\frac16, -\frac1{12})$ is a local maximum.
	\end{solution}

	\newpage
	\question[20]
	Find the maximum and the minimum of the function
	\[
		f(x,y) = x^2 + 2x + y^2 - 6y
	\]
	in the region given by 
	\[
		g(x,y) = x^2 + y^2 \le 40.
	\]
	Show your work carefully in this problem, and let us know what you are doing.
	\begin{solution}
		Use Lagrange multipliers with $g(x,y) = x^2 + y^2 = k$, for $k \in [0, 40]$.
		\begin{align*}
			\nabla f &= \vect{2x + 2, 2y-6}\\
			\nabla g &= \vect{2x, 2y}
		\end{align*}
		so we solve the system of equations
		\[
			\left\{
				\begin{aligned}
					2x+2 &= 2\lambda x \\ 
					2y-6 &= 2\lambda y \\ 
					x^2 + y^2 &= k
				\end{aligned}
			\right.
		\]
		Solving the first two equations for $\lambda$ yields
		\[
			x = \frac{1}{\lambda - 1} \qquad y = \frac{-3}{\lambda-1}
		\]
		Substituting into the third gives 
		\[
			\frac{10}{(\lambda-1)^2} = k \implies \frac{1}{\lambda-1} = \pm \sqrt{\frac{k}{10}}
		\]
		so for a fixed $k$, the optimization problem yields two solutions:
		\[
			(x, y) = \left( \pm \sqrt{\frac{k}{10}}, \mp 3 \sqrt{\frac{k}{10}} \right)
		\]
		To determine max/min values, we plug these points into $f$:
		\begin{align*}
			f &= x^2 + y^2 + 2x - 6y \\
			&= k + 2\left(\pm \sqrt{\frac{k}{10}}\right) - 6\left( \mp 3 \sqrt{\frac{k}{10}} \right)  \\
			&= k \pm 20 \sqrt{\frac{k}{10}} \\
			&= k \pm 2 \sqrt{10}\sqrt{k}
		\end{align*}
		We want to optimize $f$ with respect to $k \in [0,40]$, which requires us to evaluate $k$ at critical points and at end points.
		Note there are two cases for $f$. 
		\begin{outline}
			\1 If $f = k + 2 \sqrt{10} \sqrt{k}$, then $f' = 1 + \frac{\sqrt{10}}{\sqrt k} = 0$ which yields no critical points. 
			\2 Evaluating at $k = 0$, $f = 0$
			\2 Evaluating at $k = 40$, $f = 40 + 2 \sqrt{400} = 80$
			\1 If $f = k -2 \sqrt{10} \sqrt{k}$, then $f' = 1 - \frac{\sqrt{10}}{\sqrt{k}} = 0 \implies k = 10$ is a critical point.
			\2 Evaluating at $k = 0$, $f = 0$
			\2 Evaluating at $k = 10$, $f = 10 - 2\sqrt{100} = -10$
			\2 Evaluating at $k = 40$, $f = 40 - 2 \sqrt{400} = 0$
		\end{outline}
		Thus the maximum for this function in the specified region is 80 (which occurs at $(2, -6)$) and the minimum is $-10$ (which occurs at $(-1, 3)$).
	\end{solution}

	\newpage
	\question[8]
	Suppose that $x = r \cos \theta$ and $y = r \sin \theta$ (the usual polar coordinates) and $f(x,y) = x^2 y$. 
	Express
	\[
		\pdv{f}{r} \qand \pdv{f}{\theta}
	\]
	as functions of $r$ and $\theta$. (Hint/Comment: Do this however you like.)
	\begin{solution}
		\begin{align*}
			f &= x^2 y = r^3 \cos^2 \theta \sin \theta \\ 
			f_r &= 3r^2 \cos^2 \theta \sin\theta \\ 
			f_\theta &= r^3 \left(2\cos\theta(-\sin\theta)\sin\theta + \cos^2\theta\cos\theta \right) \\ 
			&= r^3 (-2\cos\theta\sin^2\theta + \cos^3\theta)
		\end{align*}
	\end{solution}

	\newpage
	\question[18]
	Short answers \ldots
	\begin{parts}
		\part If $f$ is a function of $x$ and $y$, and $x$ and $y$ are each functions of $r$, $s$, and $t$, then use the chain rule to express $\dpd{f}{s}$.
		\begin{solution}
			\[
				\dpd fs = \dpd fx \dpd xs + \dpd fy \dpd ys
			\]
		\end{solution}

		\part Find the average value of the function $f(x,y) = x^2y$ on the rectangle $0 \le x \le 3$, $0 \le y\le 4$.
		\begin{solution}
			\[
				\frac1{12} \int_0^3 \int_0^4 x^2y \dif y \dif x
				= \frac1{12} \left[\frac{x^3}{3}\right]_0^3 \left[\frac{y^2}{2}\right]_0^4
				= \frac1{12} (9)(8) = 6
			\]
		\end{solution}

		\part According to the theorem that we learned, what should you require of a set $S$ to guarantee that any continuous function $f$ will attain an absolute maximum and an absolute minimum on $S$?
		\begin{solution}
			The set $S$ must be closed and bounded.
		\end{solution}

		\part For the set $5x^2 + y^3 + 2z^6 - 3xy^2z^2 = 11$ write down the tangent plane at the point $(-1, -2, 1)$.
		\begin{solution}
			Letting $f = 5x^2 + y^3 + 2z^6 - 3xy^2z^2 - 11$, 
			\begin{align*}
				\nabla f &= \vect{10x - 3y^2z^2, 3y^2 - 6xyz^2, 12z^5 - 6xy^2z} \\ 
				\nabla f(-1, -2, 1) &= \vect{-22,0,36}
			\end{align*}
			\[
				\boxed{0 = -22(x+1) + 0(y+2) + 36(z-1)}
			\]
		\end{solution}

	\end{parts}
\end{questions}
\end{document}
