 \documentclass[12pt]{article}
\pagestyle{plain}

\usepackage{array, amssymb}
\usepackage{amsfonts, geometry}

\extrarowheight=5pt
\voffset=-20pt
\textheight=9.5in
%\topmargin=2in
\usepackage{mathtools}
%\usepackage{amsthm}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{stackengine}
\stackMath
\newcommand{\erf}{\operatorname{erf}}

\newcommand{\kk}{{\mathbf k}}
\newcommand{\CC}{{\mathbb C}}
\newcommand{\II}{{\mathbb I}}
\newcommand{\RR}{{\mathbb R}}
\newcommand{\QQ}{{\mathbb Q}}
\newcommand{\ZZ}{{\mathbb Z}}

\newcommand{\homm}{{\mathrm{Hom}}}


\DeclareMathOperator{\Hol}{Hol}
\DeclareMathOperator{\res}{Res}

\def\diam{\qopname\relax o{diam}}
\def\spt{\qopname\relax o{spt}}
\def\dist{\qopname\relax o{dist}}
\def\interior{\qopname\relax o{int}}
\def\supp{\qopname\relax o{supp}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Res}[1]{\stackunder{\res}{\scriptstyle #1}}
 \newcommand{\DD}{\mathbb{D}}
\newcommand{\re}{\mathfrak{Re}}
\newcommand{\im}{\mathfrak{Im}}
\parindent=0pt
\begin{document}

\begin{center} {\textbf{Analysis QE I Exam \,  June 2020}} \end{center}
{\bf Instructions:}  Do not write your name on any of the pages.  This is a closed book exam: no books, no notes, no calculators etc. Only plain papers and pens should be on your table.

\pagestyle{empty}

    \vspace{.2cm}
 
 








\begin{enumerate}[itemsep=0pt, leftmargin=.5cm, labelindent=.5in]
    
    \item 
Let $f:\R^2\to\R$ be defined by
\[
f(x,y)=\begin{cases}
(x^2+y^2)\sin\frac{1}{\sqrt{x^2+y^2}},\,\, (x,y)\neq (0,0), \\
0,\quad \quad (x,y)= (0,0).
\end{cases}\]
\begin{itemize}
\item[(a)] Is $f$ differentiable at $(0,0)$?

\item[(b)] Are the partial derivatives $D_i f$, $i=1,2$  continuous at $(0,0)$?
\end{itemize}


\bigskip
\item  Let the sequence of functions be defined by $f_n(x)=nxe^{-nx}$ on $[0,+\infty)$. Determine the pointwise limit on the given interval (if it exists) and an interval on which the convergence is uniform (if any).

Does the sequence of derivatives $(f_n')$ converge uniformly on 
 $[0,+\infty)$?

\item  Let  $f:D\to\mathbb{R}$, where  $D\subset \mathbb{R}$ is measurable. Show that $f$ is measurable if and only if the function  $g:\mathbb{R}\to \mathbb{R}$ is measurable, where $g(x)=f(x)$ for $x\in D$ and $g(x)=0$ otherwise.

\item  Let $f:\CC\longrightarrow\CC$ be given by
$$f(z)=z|z|.$$
Where is $f^\prime(z)$ defined? Where is $f(z)$ analytic?

\item Construct a map that maps the half-strip
$$S=\{z: |\re(z)|<1,\im(z)>0\}$$
conformally onto the open unit disk
$$\DD=\{z:|z|<1\}.$$

\item Let $f(z)$  be analytic in the punctured disk
$$D=\{z:|z|<1,z\not=1/2\}.$$
Suppose that $f(z)$ has a simple pole at $z=1/2$ and that 
$$\Res{z=1/2}f(z)=1.$$
Determine the coefficient $a_{-2}$ in the Laurent expansion
$$f(z)=\sum_{n=-\infty}^\infty a_n z^n,\quad 1/2<|z|<1.$$


    
    \end{enumerate}
  \end{document}
